---
title: "clase_3_diplomado"
author: "Luis Daniel Chavarría"
date: "24/1/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Clase 3 - Diplomado en analítica

```{r}
library(tidyverse)
library(lubridate)
library(naniar)
```

# Data reading & parsing

```{r}
raw_diplomate <- read_delim("../../database/clean_diplomate.csv",
                            delim = ";",
                            col_types = cols(atencion = col_character(),
                                             id = col_character(),
                                             edad = col_number(),
                                             genero = col_factor(),
                                             cronico = col_factor(),
                                             segcronico = col_factor(),
                                             eps = col_factor(),
                                             urg = col_factor(),
                                             tipurg = col_factor(),
                                             inter = col_factor(),
                                             ordurg = col_factor(),
                                             cirugia = col_factor(),
                                             proc = col_factor(),
                                             pisodes = col_factor(),
                                             esp = col_factor()
                                             ))

glimpse(raw_diplomate)
```

# Further cleaning

It's not advised to run the cleaning process everytime, the NA replacing will take a long time to compute as there are over 3 million data points.

```{r}
diplomate <- raw_diplomate %>%
  mutate(fechaing = as.POSIXct(fechaing, format = "%d/%m/%Y %H:%M:%S %p"), #The numbers are 43101,02954 (Hours)
         horaing = str_extract_all(horaing, pattern = "\\d\\d:\\d\\d:\\d\\d", simplify = TRUE),
         horaing = hms(horaing, quiet = TRUE),
         anoing = as.character(anoing),
         anoing = parse_date(anoing, format = "%Y")
         ) %>% 
  replace_with_na_all(condition = ~.x %in% common_na_strings)

# write_excel_csv2(diplomate, "diplomate_ok_na.csv")
# write_rds(diplomate, "diplomate_ok_na.rds")
glimpse(diplomate)
```

# Cleaning after NA processing

The file has been saved in a rds format, so that it differentiates correctly from the other data and loads the cleaned up formats. 

```{r}
dipl_naok <- read_rds("database/diplomate_ok_na.rds")

dipl_naok %>% mutate(
    atencion = as.character(atencion),
    id = as.character(id),
    edad = as.numeric(edad),
    genero = as_factor(genero),
    cronico = as_factor(cronico),
    segcronico = as_factor(segcronico),
    eps = as_factor(eps),
    urg = as_factor(urg),
    tipurg = as_factor(tipurg),
    inter = as_factor(inter),
    ordurg = as_factor(ordurg),
    cirugia = as_factor(cirugia),
    proc = as_factor(proc),
    pisodes = as_factor(pisodes),
    esp = as_factor(esp))
```

