---
title: "Data cleaning new diplomate"
author: "Daniel"
date: "27/2/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries

```{r}
library(tidyverse)
library(naniar)
library(readxl)
library(lubridate)
library(writexl)
```

# Data reading

raw_data2 includes correct parsing from Excel with dates read from 1899-12-30 but generates NA for the data in different format, whereas raw_data coerces everything to character and keeps the information intact.

```{r}
raw_data <- readxl::read_excel("database/Datos_diplomado R.xlsx", 
                                col_types = c("text", "text", "text", 
                                "text", "numeric", "text", "text", 
                                "text", "text", "numeric", "text", 
                                "numeric", "text", "text", "text", 
                                "text", "text", "text", "text", "text", 
                                "text", "text", "numeric", "numeric", 
                                "numeric", "numeric", "numeric", 
                                "numeric", "text", "text", "text", 
                                "text", "text", "text", "text", "numeric", 
                                "numeric", "numeric", "text", "numeric"))

raw_data2 <- readxl::read_excel("database/Datos_diplomado R.xlsx", 
                                col_types = c("text", "text", "text", 
                                "date", "numeric", "text", "text", 
                                "text", "date", "numeric", "text", 
                                "numeric", "text", "text", "text", 
                                "text", "text", "text", "text", "text", 
                                "text", "text", "numeric", "numeric", 
                                "numeric", "numeric", "numeric", 
                                "numeric", "text", "text", "text", 
                                "text", "numeric", "text", "numeric", "numeric", 
                                "numeric", "numeric", "numeric", "numeric"))
```

# fecha_nacimiento

Which dates are in a different format in fecha_nacimiento?

```{r}
raw_data %>% 
  count(fecha_nacimiento, sort = T) %>% distinct(fecha_nacimiento, .keep_all = T) %>%
  mutate(str_size = stringr::str_length(fecha_nacimiento)) %>% 
  filter(str_size > 5) %>% 
  inner_join(raw_data %>%
               distinct(fecha_nacimiento, .keep_all = T) %>% 
               dplyr::select(Atencion, fecha_nacimiento))
```

So, we have dd/mm/aaaa and dd Feb aaaa, let's fix that. `dmy()` can handle both "20 feb 1997" and "25/4/1972"

Why is the representation of 27/06/1930 26894, because Excel has origin at 1899-12-30 and counts days from that date.

```{r}
dmy("20 feb 1997") #For proof
dmy("25/4/1972") #For proof
daymonthyear <- dmy(raw_data$fecha_nacimiento)

correct_nacimiento <- raw_data2
correct_nacimiento$fecha_nacimiento[is.na(raw_data2$fecha_nacimiento)] <- daymonthyear[is.na(raw_data2$fecha_nacimiento)]
```

Cleaned dates for `fecha_nacimiento` are now stored correctly in `correct_nacimiento`

# fecha_atenc

```{r}
correct_attention <- raw_data2 %>%
  mutate(mes_atencion = case_when(mes_atencion == "1/01/febrero000" ~ "octubre",
                                  mes_atencion == "1/01/febrero00febrero" ~ "octubre",
                                  TRUE ~ mes_atencion)) %>%
  mutate(fecha_atenc = case_when(is.na(fecha_atenc) ~ ymd(paste(ano_atencion, mes_atencion, dia_atencion)),
                                 TRUE ~ as_date(as.character(fecha_atenc))))
  
```

Las 57 atenciones marcadas como NA ocurren porque según la información, dice que ocurrió el 29 de enero de 2018, y esa fecha no existe dado que no fue un año bisiesto y no existe tal combinación. Se analizan los datos y se identifica que los 1/01/febrero000 corresponden a octubre.

# Otras correciones

- Se añaden ambas fechas correctamente
- Se elimina la columna identificación
- Se reemplazan "NA" string a "no aplica"

```{r}
diplomado_clean <- correct_attention %>%
  mutate(fecha_nacimiento = correct_nacimiento$fecha_nacimiento) %>% 
  select(-identificacion) %>%
  mutate(fecha_nacimiento = as_date(fecha_nacimiento)) %>%
  mutate_if(.predicate = is.character, .funs = funs(str_replace(., "NA", "no aplica"))) %>% 
  mutate(nombres = str_replace(nombres, "no aplica", "NA"))
```

# Escritura

```{r}
writexl::write_xlsx(diplomado_clean, "database/datos_diplomado_limpios.xlsx")
```
